---
title: "Vaccination Data Analysis"
author: "Mohan Teja,"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading data

```{r}
library(readr)
vaccinations <- read_csv("data/vaccinations.csv",show_col_types = FALSE)
newCases <- read_csv("data/new_cases_per_million.csv",show_col_types = FALSE)
deaths <- read_csv("data/total_deaths_per_million.csv",show_col_types = FALSE)
```

## Subsetting the Data

### Note : 

new cases and deaths are taken per million and vaccinations are taken per 100

```{r}
ind <-vaccinations[vaccinations[1]=="India",]
vaxIndia <- data.frame(ind$date,ind$people_fully_vaccinated_per_hundred)
summary(vaxIndia)
newCasesIndia <- data.frame(newCases$date,newCases$India)
summary(newCasesIndia)
deathsIndia <- data.frame(deaths$date,deaths$India)
summary(deathsIndia)
```



## Cleaning Data 

### Na values

As we can see there are Na values in the data and timelines don't algin correctly we can fill na with previous know values since this is a time series data so we fill na with previous known values and remaining na with 0

```{r}
fill.NAs <- function(x) {is_na<-is.na(x); x[Reduce(function(i,j) if (is_na[j]) i else j, seq_len(length(x)), accumulate=T)]}
vaxIndia$ind.people_fully_vaccinated_per_hundred <- fill.NAs(vaxIndia$ind.people_fully_vaccinated_per_hundred)
vaxIndia[is.na(vaxIndia)] = 0
summary(vaxIndia)
newCasesIndia$newCases.India <- fill.NAs(newCasesIndia$newCases.India)
newCasesIndia[is.na(newCasesIndia)] = 0 
summary(newCasesIndia)
deathsIndia$deaths.India <- fill.NAs(deathsIndia$deaths.India)
deathsIndia[is.na(deathsIndia)] = 0
summary(deathsIndia)
```

### Dates

```{r}
vaxIndia$ind.date = as.Date(vaxIndia$ind.date,"%y-%m-%d")
newCasesIndia$newCases.date = as.Date(newCasesIndia$newCases.date,"%y-%m-%d")
deathsIndia$deaths.date = as.Date(deathsIndia$deaths.date,"%y-%m-%d")

vaxIndia  = vaxIndia[vaxIndia$ind.date > as.Date("2021-01-14") & vaxIndia$ind.date < as.Date("2022-03-02"),]
summary(vaxIndia)
newCasesIndia = newCasesIndia[newCasesIndia$newCases.date > as.Date("2021-01-14") & newCasesIndia$newCases.date < as.Date("2022-03-02"),]
summary(newCasesIndia)
deathsIndia = deathsIndia[deathsIndia$deaths.date > as.Date("2021-01-14") & deathsIndia$deaths.date < as.Date("2022-03-02"),]
summary(deathsIndia)
```

Number of deaths is now cumulative data we need to make it non cumulative and make a data frame out of all the clean lists we made

```{r}
ndeaths = diff(c(0,deathsIndia$deaths.India))
ndeaths[1]=0
df = data.frame(ncases = newCasesIndia$newCases.India,vax = vaxIndia$ind.people_fully_vaccinated_per_hundred,death = ndeaths)
summary(df)
```

## Correlations

```{r}
c =  cor(df)
corrplot::corrplot.mixed(c,order='AOE')
```

### Observation :

* We can see that vaccinations have little to no effect on number of new cases  
* New cases have a very high positive correlation with deaths
* Vaccinations have a negative correalation with death

## plots
```{r}
plot.ts(df)
```

## Model
```{r}
mdl = lm(death~vax*ncases,data = df)
summary(mdl)
d =  predict(mdl,newdata = df)
plot(df$death,type = "l")
lines(d,col="red")
```

### Result :
Hence we can see that reducing new cases and increasing vaccination will increase mortality ie decrease number of new deaths